{ \fontfamily{ptm}\selectfont First Order Nonsmooth Optimization: Algorithm Design, Variational analysis, and Applications }

### 2025-01-10 Notebook LaTEX/Proposal/main.tex
>w<, todonotes removed

\section{Introduction}
    Let $\RR^n$ be the ambient space. We consider 
    \begin{align}
        \min_{x \in \RR^n} \left\lbrace
            F(x): f(x) + g(x)
        \right\rbrace.
    \end{align}\label{eqn:additive-comp-obj}
    Unless specified, assume $f:\RR^n \rightarrow \RR$ is $L$-Lipscthiz smooth $\mu \ge 0$ strongly convex and $g:Q \rightarrow \overline \RR$ is convex. 
    This type of problem is referred to as additive composite problems in the literature. 
    \par
    Our ongoing research concerns accelerated proximal gradient type method for solving (\ref{eqn:additive-comp-obj}). 
    In the expository writing by Walkington \cite{noel_nesterovs_nodate}, a variant for of accelerated gradient method for strongly convex function $f$ is discussed. 
    We had some lingering questions after reading it. 
    \begin{enumerate}
        \item Do there exist a unified description for the convergence for both variants of the algorithms?
        \item Is it possible to attain fast convergence rate without knowledge about the strong convexity of function $f$?
        \item Is it possible to describe the convergence of function value for momentum sequences that are much weaker than the Nesterov's rule? 
    \end{enumerate}
    The good news is we have definitive answers for all questions by our own efforts of research. 
    \todoinline{\noindent
        $\square$ Change all the organization parts here
    }
    Section \ref{sec:unify-nes-acceleration}, \ref{sec:free-rwapg} are our ongoing research which present the answers to the questions. 
    \par
    In Section \ref{sec:unify-nes-acceleration}, we proposed the method of ``Relaxed Weak Accelerated Proximal Gradient (R-WAPG)'' as the foundation to describe several variants of Accelerated proximal gradient method in the literatures. 
    The convergence theories of R-WAPG allows us to model convergence of accelerated proximal gradient method where the momentum sequence doesn't strictly follow the conditions presented in the literatures. 
    The descriptive power of R-WAPG allows convergence analysis for all the variants using one single theorem. 
    \par
    In Section \ref{sec:free-rwapg} we propose a practical algorithm that exploits a specific term in the proof of R-WAPG to achieve faster convergence for solving (\ref{eqn:additive-comp-obj}) without knowing parameter $L, \mu$ in prior. 
    Results of numerical experiments are presented. 
    \par
    Section \ref{sec:catalyst} complements report completed for MATH 590 Fall Winter 2024. 
    \todo{
        $\blacksquare$ Add Lin's papers and Paquette's papers.
    }
    {\hl{It's based on a series of recent papers}} \cite{lin_universal_2015,lin_catalyst_2018,paquette_catalyst_2018} about Catalyst Meta Acceleration for First Order Variance Reduced Methods. 
    The end of Section \ref{sec:catalyst} points out potential future direction of research of Catalyst Acceleration. 
    % \par
    % Section \ref{sec:inexact-prox}, \ref{sec:nes-acc-ncnvx} preview literatures in nonsmooth optimization frontier research where progress and impacts can be made.  
    \subsection{Theme of the research}
        This section specifies a theme of the research in this proposal. 
        Firstly, we explore the Goldilocks zones between these topics: Nesterov's acceleration, algorithms with inexact evaluation of the proximal point operator, and applications.  
        In doing so, we identify the ``chemistry'' between various properties of functions and the affect it has on the designs/behaviors of first order algorithm for continuous optimization.

@@

            \todo{$\blacksquare$ Add bib \\ $\blacksquare$ Cite. }
            \hl{Pang, Cui's book} \cite{ying_modern_2021}

@@

            \todo{$\blacksquare$ Add bib \\ $\blacksquare$ Cite. }
            {\hl{Pang, Cui's book}} \cite{ying_modern_2021}

@@

    \todo{
        $\blacksquare$ Add Nesterov's original paper. \\ 
        $\blacksquare$ Cite it.  
    {\hl{Proposed back in 1983 the original Nesterov's acceleration method}} \cite{nesterov_method_1983}

@@

    \todo{
        $\blacksquare$ Add Nesterov's new book. 
        \\ $\blacksquare$ Cite it.
    {\hl{This convergence rate is considered optimal for all class of $L$ Lipschitz smooth convex function }}\cite{nesterov_lectures_2018}. 

@@

    \todo{
        $\blacksquare$ Add Beck 2009 FISTA original paper. \\
        $\blacksquare$ Cite it.
    {\hl{Beck and Toubelle}} \cite{beck_fast_2009}

@@

    \todo{
        $\blacksquare$ Add Necoara linear convergence 
        $\blacksquare$ Add Aujol 2024 Parameter free FISTA restart. 
        $\blacksquare$ Cite Section 5.2.2 for the former, cite the entirety for the latter. 
    {\hl{See} (5.2.2) Necoara et al. \cite{necoara_linear_2019} and Aujol et al. \cite{aujol_parameter-free_2024}}

@@

    \todo{
        $\blacksquare$ Add Neol J. Walkington's ``Nesterov's Method for Convex  Optimization''. \\ 
        $\blacksquare$ Cite it. 
    {\hl{Walkington's writing}} \cite{noel_nesterovs_nodate}, 

@@

    \todo{
        $\blacksquare$ Add Beck's first order textbook
        \\ $\blacksquare$ Cite 10.7.7. 
        \\ $\blacksquare$ Add Nesterov's textbook. 
        \\ $\blacksquare$ Cite it. }
    {\hl{
        The same variant also appears in Beck's book as V-FISTA} \cite[(10.7.7)]{beck_first-order_2017}, and \hl{Nesterov's book as (2.2.22)} \cite{nesterov_lectures_2018}.

@@

    \todo{
        $\blacksquare$ Add paper: ``On the convergence of the iterates of the...''
        $\blacksquare$ Cite them. }
    {\hl{Chambolle, Dossal}} \cite{chambolle_convergence_2015}

@@

        \todo{
            $\blacksquare$ Add ``On the convergence of the iterates of...'' \\ $\blacksquare$ Cite them.}
        {\hl{Chambolle, Dossal}} \cite{chambolle_convergence_2015},
        and we are able to show the same convergence rate in Theorem \ref{thm:r-wapg-on-cham-doss}. 
        When $\rho_k = 1, \mu = 0$, R-WAPG reduces perfectly to FISTA by in 
        \todo{
            $\blacksquare$ Add beck's original FISTA Paper. 
            \\$\blacksquare$ Cite it.
        {\hl{Beck}} \cite{beck_fast_2009}.

@@

        Organization now follows. 
        Section \ref{ssec:building-block-rwapg} provides a stepwise description of the R-WAPG iterative algorithm along with an inequality crucial to proving the convergence rate later. 
        Section \ref{ssec:rwpag-seq-and-alg} introduce the definition of an R-WAPG sequence, which constraints all possible parameters used permitted by the algorithm.
        The section also states the full R-WAPG algorithm and an upper bound on $F(x_k) - F(x^*)$. 
        Here, $x_k$ is generated by the R-WAPG algorithm and $x^*$ is the minimizer. 
        Section \ref{ssec:quiv-repr-rwapg} bring forward three equivalent forms of the R-WAPG algorithm making it more comparable with other Accelerated Proximal Gradient method appeared in the literatures. 
        Section \ref{ssec:describe-variants-with-rwapg} gives characterizations of specific R-WAPG sequence that leads to convergence of the R-WAPG algorithm in terms of the optimality gap. 
        The section also identifies specific instance of permissible R-WAPG sequences where it fits with the FISTA, V-FISTA, and the algorithm proposed by 
        \todo{
            $\blacksquare$ Cite the Chambolle, Dossal 2015 paper here again. 
        }
        {\hl{Chambolle Dossal}} \cite{chambolle_convergence_2015}. 
        Section \ref{sec:free-rwapg} provides Algorithm \ref{alg:free-rwapg} (Free R-WAPG) which is a parameter free alternative to accelerated proximal gradient method formulated using the R-WAPG we proposed. 
        The algorithm dynamically adjusts the R-WAPG sequence $\alpha_k$ using the Bregman Divergence of smooth part of the objective function at successive extrapolated iterate $y_{k + 1}, y_k$ from the momentum method. 
        Numerical experiments are conducted repeatedly and statistics on the results are displayed.

    \subsection{Building Blocks of R-WAPG}\label{ssec:building-block-rwapg}
        Definition\ref{def:stepwise-wapg} describes the procedures of generating the iterates $(v_{k + 1}, x_{k + 1})$ given any $(v_k, x_k)$ and parameter $\alpha_k \in (0, 1), \gamma_k > 0$. 
        Proposition \ref{prop:stepwise-lyapunov} gives an inequality instrumental to the convergence rate analysis, with the same assumption as the definition.

@@

        \todo{$\blacksquare$ Add and explain this part. }{\hl{

@@

        }}

@@

            \todo{
                $\blacksquare$ Add them \\ 
                $\blacksquare$ Cite them. 
            {\hl{in Ryu and Yin's Book} \cite{ryu_large-scale_2022}}, 
            right after Theorem 17. 

@@

            Finally, a similar triangle for formulation of FISTA can be found in Equation (2), (3), (4) in 
            \todo{$\blacksquare$ Cite Chambolle Dossal 2015 again. }
            \cite{chambolle_convergence_2015}. 

@@

            \todo{
                $\blacksquare$ Add these. $\blacksquare$ Cite these. 
            {(3.1, 4.1) in Lee et al. \cite{lee_geometric_2021} and Ahn and Sra \cite{ahn_understanding_2022}}. 

@@

    \subsection{The descriptive power of R-WAPG on existing variants}\label{ssec:describe-variants-with-rwapg}
        Lemma \ref{lemma:inverted-fista-seq} identifies a sequence $(\alpha_k)_{k \ge 0}$ such that $\alpha_k^{-2} \ge \alpha_{k + 1}^{-2} - \alpha_{k + 1}^{-1}$ as a specific instance of R-WAPG sequence. 
        The lemma showed that sequence $(\alpha_k^{-1})_{k \ge 0}$ is the FISTA sequence which governs the momentum term and convergence claim in FISTA algorithms and variants alike. 
        The lemma also provides a simplified convergence claim using the R-WAPG sequence on Proposition \ref{prop:wagp-convergence}. 
        Theorem \ref{thm:r-wapg-on-cham-doss} stated that the sequences given in  
        \todo{$\blacksquare$ Cite this paper again here. }
        {\hl{Chambolle, Dossal's}}
        paper \cite{chambolle_convergence_2015} indeed is an instance of R-WAPG sequence, along with that, it indeed attains a convergence rate $\mathcal O(1/k^2)$. 
        \par

@@

        \todo{
            $\blacksquare$ Add reference ``Regression shrinkage and selection via the Lasso''
            \\ $\blacksquare$ Cite it. 
        }{\hl{Tibshirani}} \cite{tibshirani_regression_1996}. 

>>>>

\section{Introduction}
    Let $\RR^n$ be the ambient space. We consider 
    \begin{align}
        \min_{x \in \RR^n} \left\lbrace
            F(x): f(x) + g(x)
        \right\rbrace.
    \end{align}\label{eqn:additive-comp-obj}
    Unless specified, assume $f:\RR^n \rightarrow \RR$ is $L$-Lipscthiz smooth $\mu \ge 0$ strongly convex and $g:Q \rightarrow \overline \RR$ is convex. 
    This type of problem is referred to as additive composite problems in the literature. 
    \par
    Our ongoing research concerns accelerated proximal gradient type method for solving (\ref{eqn:additive-comp-obj}). 
    In the expository writing by Walkington \cite{noel_nesterovs_nodate}, a variant for of accelerated gradient method for strongly convex function $f$ is discussed. 
    We had some lingering questions after reading it. 
    \begin{enumerate}
        \item Do there exist a unified description for the convergence for both variants of the algorithms?
        \item Is it possible to attain fast convergence rate without knowledge about the strong convexity of function $f$?
        \item Is it possible to describe the convergence of function value for momentum sequences that are much weaker than the Nesterov's rule? 
    \end{enumerate}
    The good news is we have definitive answers for all questions by our own efforts of research. 
    \todoinline{\noindent
        $\square$ Change all the organization parts here
    }
    Section \ref{sec:unify-nes-acceleration}, \ref{sec:free-rwapg} are our ongoing research which present the answers to the questions. 
    \par
    In Section \ref{sec:unify-nes-acceleration}, we proposed the method of ``Relaxed Weak Accelerated Proximal Gradient (R-WAPG)'' as the foundation to describe several variants of Accelerated proximal gradient method in the literatures. 
    The convergence theories of R-WAPG allows us to model convergence of accelerated proximal gradient method where the momentum sequence doesn't strictly follow the conditions presented in the literatures. 
    The descriptive power of R-WAPG allows convergence analysis for all the variants using one single theorem. 
    \par
    In Section \ref{sec:free-rwapg} we propose a practical algorithm that exploits a specific term in the proof of R-WAPG to achieve faster convergence for solving (\ref{eqn:additive-comp-obj}) without knowing parameter $L, \mu$ in prior. 
    Results of numerical experiments are presented. 
    \par
    Section \ref{sec:catalyst} complements report completed for MATH 590 Fall Winter 2024. 
    It's based on a series of recent papers \cite{lin_universal_2015,lin_catalyst_2018,paquette_catalyst_2018} about Catalyst Meta Acceleration for First Order Variance Reduced Methods. 
    The end of Section \ref{sec:catalyst} points out potential future direction of research of Catalyst Acceleration. 
    % \par
    % Section \ref{sec:inexact-prox}, \ref{sec:nes-acc-ncnvx} preview literatures in nonsmooth optimization frontier research where progress and impacts can be made.  
    \subsection{Theme of the research}
        This section specifies a theme of the research in this proposal. 
        Firstly, we explore the Goldilocks zones between these topics: Nesterov's acceleration, algorithms with inexact evaluation of the proximal point operator, and applications.  
        In doing so, we identify the ``chemistry'' between various properties of functions and the affect it has on the designs/behaviors of first order algorithm for continuous optimization.

@@

            Pang, Cui's book \cite{ying_modern_2021}

@@

            Pang, Cui's book \cite{ying_modern_2021}

@@

    Proposed back in 1983 the original Nesterov's acceleration method \cite{nesterov_method_1983}

@@

    This convergence rate is considered optimal for all class of $L$ Lipschitz smooth convex function \cite{nesterov_lectures_2018}. 

@@

    Beck and Toubelle \cite{beck_fast_2009}

@@

    See (5.2.2) Necoara et al. \cite{necoara_linear_2019} and Aujol et al. \cite{aujol_parameter-free_2024}

@@

    Walkington's writing \cite{noel_nesterovs_nodate}, 

@@

    The same variant also appears in Beck's book as V-FISTA \cite[(10.7.7)]{beck_first-order_2017}, and \hl{Nesterov's book as (2.2.22)} \cite{nesterov_lectures_2018}.

@@

    Chambolle, Dossal \cite{chambolle_convergence_2015}

@@

        Chambolle, Dossal \cite{chambolle_convergence_2015}, and we are able to show the same convergence rate in Theorem \ref{thm:r-wapg-on-cham-doss}. 
        When $\rho_k = 1, \mu = 0$, R-WAPG reduces perfectly to FISTA by in Beck \cite{beck_fast_2009}.

@@

        Organization now follows. 
        Section \ref{ssec:building-block-rwapg} provides a stepwise description of the R-WAPG iterative algorithm along with an inequality crucial to proving the convergence rate later. 
        Section \ref{ssec:rwpag-seq-and-alg} introduce the definition of an R-WAPG sequence, which constraints all possible parameters used permitted by the algorithm.
        The section also states the full R-WAPG algorithm and an upper bound on $F(x_k) - F(x^*)$. 
        Here, $x_k$ is generated by the R-WAPG algorithm and $x^*$ is the minimizer. 
        Section \ref{ssec:quiv-repr-rwapg} bring forward three equivalent forms of the R-WAPG algorithm making it more comparable with other Accelerated Proximal Gradient method appeared in the literatures. 
        Section \ref{ssec:describe-variants-with-rwapg} gives characterizations of specific R-WAPG sequence that leads to convergence of the R-WAPG algorithm in terms of the optimality gap. 
        The section also identifies specific instance of permissible R-WAPG sequences where it fits with the FISTA, V-FISTA, and the algorithm proposed by Chambolle Dossal \cite{chambolle_convergence_2015}. 
        Section \ref{sec:free-rwapg} provides Algorithm \ref{alg:free-rwapg} (Free R-WAPG) which is a parameter free alternative to accelerated proximal gradient method formulated using the R-WAPG we proposed. 
        The algorithm dynamically adjusts the R-WAPG sequence $\alpha_k$ using the Bregman Divergence of smooth part of the objective function at successive extrapolated iterate $y_{k + 1}, y_k$ from the momentum method. 
        Numerical experiments are conducted repeatedly and statistics on the results are displayed.

    \subsection{Building Blocks of R-WAPG}\label{ssec:building-block-rwapg}
        Definition \ref{def:stepwise-wapg} describes the procedures of generating the iterates $(v_{k + 1}, x_{k + 1})$ given any $(v_k, x_k)$ and parameter $\alpha_k \in (0, 1), \gamma_k > 0$. 
        Proposition \ref{prop:stepwise-lyapunov} gives an inequality instrumental to the convergence rate analysis, with the same assumption as the definition.

@@

@@

@@

            in Ryu and Yin's Book \cite{ryu_large-scale_2022}, right after Theorem 17. 

@@

            Finally, a similar triangle for formulation of FISTA can be found in Equation (2), (3), (4) in \cite{chambolle_convergence_2015}. 

@@

            (3.1, 4.1) in Lee et al. \cite{lee_geometric_2021} and Ahn and Sra \cite{ahn_understanding_2022}. 

@@

    \subsection{The descriptive power of R-WAPG on existing variants}\label{ssec:describe-variants-with-rwapg}
        Lemma \ref{lemma:inverted-fista-seq} identifies a sequence $(\alpha_k)_{k \ge 0}$ such that $\alpha_k^{-2} \ge \alpha_{k + 1}^{-2} - \alpha_{k + 1}^{-1}$ as a specific instance of R-WAPG sequence. 
        The lemma showed that sequence $(\alpha_k^{-1})_{k \ge 0}$ is the FISTA sequence which governs the momentum term and convergence claim in FISTA algorithms and variants alike. 
        The lemma also provides a simplified convergence claim using the R-WAPG sequence on Proposition \ref{prop:wagp-convergence}. 
        Theorem \ref{thm:r-wapg-on-cham-doss} stated that the sequences given in Chambolle, Dossal's
        paper \cite{chambolle_convergence_2015} indeed is an instance of R-WAPG sequence, along with that, it indeed attains a convergence rate $\mathcal O(1/k^2)$. 

@@

        Tibshirani \cite{tibshirani_regression_1996}. 
